---
title: "Target_People"
author: "Sarah Deussing"
date: "2025-04-17"
output: html_document
---

Target group qualifications:

  1. Age >= 25 and Age < 45
  2. Increase potential consecutive years (projected time as a donor) or increase gift amount
      - Factors that increase consecutive years: Active Giving Society, Monogram, Dual Domer Household
      - Factors that increase gift amount: Active Giving Society, Current Parent

Reasoning:

  1. Eliminate the dropoff in donors
  2. From models, those factors are shown to increase.

Requirements:

  1. 0.25% of people must be >= 70.5 
  2. 1,000 from each region
  
  
```{r, message = FALSE, warning = FALSE}
library(dplyr)
old_portfolio = read.csv("Old Portfolio.csv")

# cleaning
# Char --> Date
cols1_2y <- c("Most.Recent.UNR.Gift.Date", "Sorin...Renewal.Date", "Rockne...Renewal.Date", "Law...Renewal.Date", "Grad.Bus...Renewal.Date")
cols1_4y <- c("Most.Recent.Visit.Date")

old_portfolio[cols1_2y] <- lapply(old_portfolio[cols1_2y], function(x) as.Date(x, format = "%m/%d/%y"))
old_portfolio[cols1_2y] <- lapply(old_portfolio[cols1_4y], function(x) as.Date(x, format = "%m/%d/%Y"))

old_portfolio$Last.Gift.Days.Ago = as.numeric(Sys.Date() - old_portfolio$Most.Recent.UNR.Gift.Date)

# Take out $ at start --> numeric
cols2 <- c("Most.Recent.UNR.Gift.Legal.Amount", "Most.Recent.UNR.Gift.Credit.Amount", "Capacity.Rating...Rating.Low.Amount.Dim", "Lifetime.Production.Household.Extended.Amount", "Current.Campaign.Production.Household.Extended.Amount", "Sorin...Current.Year..", "Sorin...One.Year.Prior..", "Rockne...Current.Year..", "Rockne...One.Year.Prior..", "Law...Current.Year..", "Law...One.Year.Prior..", "Grad.Bus...Current.Year..", "Grad.Bus...One.Year.Prior..", "X2025.FY.Matching.Gifts", "X2024.FY.Matching.Gifts", "X2025.FY.Total.Unrestricted.Gifts", "X2024.FY.Total.Unrestricted.Gifts", "X2023.FY.Total.Unrestricted.Gifts", "X2022.FY.Total.Unrestricted.Gifts", "X2021.FY.Total.Unrestricted.Gifts")
old_portfolio[cols2] <- lapply(old_portfolio[cols2], function(x) as.numeric(gsub("[$,]", "", x)))

old_portfolio$Zip.Code <- as.numeric(old_portfolio$Zip.Code)

# Y/N to 1/0
cols3 <- c("X70.5..Ind", "Has.Spouse.Indicator", "Participation.Donor.Current.Fiscal.Year", "Current.Parent.Ind", "Monogram...Big.4", "Monogram...Non.Big.4", "Dual.Domer.Household")
old_portfolio[cols3] <- lapply(old_portfolio[cols3], function(x) ifelse(x == "Y", 1, ifelse(x == "N", 0, NA)))

old_portfolio <- old_portfolio %>%
  mutate(General.Active.Giving.Society = sub("^(\\S+).*", "\\1", Active.Giving.Society.Affiliation.List),
         General.Gift.Designation = sub("^(\\S+).*", "\\1", Most.Recent.UNR.Gift.Designation))

#summary(old_portfolio)
```

Set Region
```{r}
assign_region <- function(state, zip) {
  zip_prefix <- as.numeric(substr(as.character(zip), 1, 3))
  
  if (state == "CA") {
    if (!is.na(zip_prefix) && zip_prefix < 936) {
      return("Southwest")
    } else {
      return("Northwest")
    }
  } else if (state == "NV") {
    if (!is.na(zip_prefix) && zip_prefix < 898) {
      return("Southwest")
    } else {
      return("Northwest")
    }
  } else if (state == "WI") {
    if (!is.na(zip_prefix) && zip_prefix < 536) {
      return("Illinois")
    } else {
      return("Northwest")
    }
  } else if (state == "PA") {
    if (!is.na(zip_prefix) && zip_prefix < 160) {
      return("Central")
    } else {
      return("East Coast")
    }
  } else if (state %in% c("WA", "OR", "MT", "ID", "ND", "SD", "NE", "MN", "IA")) {
    return("Northwest")
  } else if (state %in% c("CO", "AZ", "UT")) {
    return("Southwest")
  } else if (state %in% c("IL")) {
    return("Illinois")
  } else if (state %in% c("MI", "IN", "OH", "KY")) {
    return("Central")
  } else if (state %in% c("TN", "GA", "WV", "VA", "NC", "MD", "DE", "VT", "NH", "MA", "CT", "RI", "ME")) {
    return("East Coast")
  } else if (state %in% c("NY", "NJ")) {
    return("NY Metro")
  } else if (state %in% c("NM", "KS", "TX", "OK", "MO", "AR", "LA", "MS", "AL", "FL")) {
    return("Southern")
  } else {
    return("Unknown")
  }
}

old_portfolio$Region <- mapply(assign_region, old_portfolio$Preferred.Address...State.Code,
                               old_portfolio$Zip.Code)
```


Age Filter
```{r}
target_age <- old_portfolio %>% filter(Age >= 25 & Age < 45)
nrow(target_age)
```

Top Donors Based on Model Characteristics
```{r}
top_donors <- old_portfolio %>% filter(Current.Parent.Ind == 1 | Dual.Domer.Household == 1 | Monogram...Big.4 == 1 | Monogram...Non.Big.4 == 1 | Active.Giving.Society.Affiliation.List != "")

nrow(top_donors)
```


```{r}
target2 <- rbind(top_donors, target_age) %>% distinct()
nrow(target2)
```


Want 1,000 per major region
```{r}
target2 <- target2 %>% filter(Region != "Unknown")
target2 %>% group_by(Region) %>%
  summarize(count = n())
```

Regions with > 100 donors: Central, East Coast, Illinois, Southern.
For those regions, take the top 1000 based on what they gave in 2024.

Regions with < 1000 donors: NY Metro, Northwest, Southwest
For those regions, add additional people based on what they gave in 2024.

```{r}
# over
over <- c("Central", "East Coast", "Illinois", "Southern")
over_regions <- target2 %>% filter(Region %in% over) %>%
  group_by(Region) %>%
  arrange(desc(X2024.FY.Total.Unrestricted.Gifts), .by_group = TRUE) %>%
  slice_head(n = 1000) %>%
  ungroup()

# under
slices <- list()
under_regions <- data.frame()
under <- c("NY Metro", "Northwest", "Southwest")
counts <- target2 %>% filter(Region %in% under) %>% count(Region, name = "n_existing")
to_fill <- c(1000, 1000, 1000) - counts$n_existing

for (i in 1:3) {
  slices[[i]] <- old_portfolio %>% filter(Region == under[i], !(ID.Number %in% target2$ID.Number)) %>%
    arrange(desc(X2024.FY.Total.Unrestricted.Gifts)) %>%
    slice_head(n = to_fill[i])
}
under_regions <- bind_rows(slices)
under_regions <- rbind(under_regions, target2 %>% filter(Region %in% under))
```

Merge
```{r}
target2 <- data.frame()
target3 <- rbind(over_regions, under_regions)

target3 %>% group_by(Region) %>%
  summarize(count = n())
```


```{r}
nrow(target3 %>% filter(X70.5..Ind == 1)) / nrow(target3)
```

need to get more older people